---
title: 'Team All Stars Lab 12: Gapminder Permutation Tests'
author: "Johnny Tamanaha, Emily Archambault, Kaiyuan Fan, Eugene Miller, Jingzhi Zhang"
date: "November 15, 2019"
output: html_document
---

## Permutation Testing
```{r, message = F, echo = T, message = F, eval = F}

perm_mean <- function(perms = 1000, all_values, n_A)
{
  ## Variables ##
  # perms: The number of permutations 
  # all_values (num): all data values
  # n_A (int): Size of group A
  ###############
  
 mu_vec <- numeric(perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    #Creates a randomized vector.
    shuffle <- sample(all_values)
    
    #Creates group A and B vectors
    r_A <- shuffle[1:n_A]
    r_B <- shuffle[(n_A+1):length(all_values)]
    
    #Calculate dealta mu
    delta_mu <- mean(r_A) - mean(r_B)
    
    mu_vec[i] <- delta_mu
    
  }
  
  return(vec)
  
}
```

## Linear Correlation Permuation Test

```{r, message = F, echo = T, message = F, eval = F}

perm_cor <- function(perms = 1000, x, y)
{
  ## Variables ##
  # perms: The number of permutations 
  # x: Vector of Variable 1 - for computing correlation
  # y: Vector of Variable 2 - for computing correlation
  ###############

  #Create zero vector
  vec <- numeric(perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    #Randomize y vector
    r_Y <- sample(y)
    
    #Computer correlation and store
    vec[i] <- cor(x, r_Y)
    
  }
  
  # Return new updated vector
  return(vec)
}
```

# Gapminder Questions

Your data science team must come up with two specific questions of interest about the gapminder data set (from 'gapminder' library). 

* Briefly describe the gapminder dataset.

* Explain why your questions are important. One of the questions should be answered with the two-sample mean permutation test, and the other with the correlation permutation test.

* Finish the function skeletons to implement the two-sample permuation test and the correlation permutation test. Run a two-sample mean permutation test. Run a correlation permutation test.

* Create visualizations of the statistical tests.

* Make data-based conclusions about the results. Do you believe there are differences among the groups you are comparing? 

* Generate percentiles, summary statistics (max, min), and probabilites (p-values) to justify your conclusions.

* Summarize each individual's contribution to the lab.

# Format and Order for Team Lab Report

Create a team R Markdown document knitted to html and submitted to OSF. Put your team name in the filename. Make sure your code works on the updated dataset realeased on Friday.