---
title: 'Team All Stars Lab 12: Gapminder Permutation Tests'
author: "Johnny Tamanaha, Emily Archambault, Kaiyuan Fan, Eugene Miller, Jingzhi Zhang"
date: "November 15, 2019"
output: html_document
---
```{r, setup, message = F, echo = F, message = F, eval = T}
library(tidyverse)
library(gapminder)
```
```{r, message = FALSE}

vowels <- gapminder %>%
  filter(continent == "Asia" | continent == "Europe") %>%
  filter(year == 2007) %>%
  select("country", "continent") %>%
  mutate(vowel = str_count(country, "[AEIOUaeiou]"))

#33 countries in Asia, 30 in Europe

#Calculating the rate of vowels in a country's name
vowels$country <- as.character(vowels$country)
vowels <- vowels %>%
  mutate(rate_vowel = vowel/nchar(`country`))
asia_vowels <- vowels %>%
  filter(continent == "Asia")
europe_vowels <- vowels %>%
  filter(continent == "Europe")
```

## Permutation Testing
```{r, message = F, echo = T, message = F, eval = T}

perm_mean <- function(perms = 1000, all_values, n_A)
{
  ## Variables ##
  # perms: The number of permutations 
  # all_values (num): all data values
  # n_A (int): Size of group A
  ###############
  
 mu_vec <- numeric(perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    #Creates a randomized vector.
    shuffle <- sample(all_values)
    
    #Creates group A and B vectors
    r_A <- shuffle[1:n_A]
    r_B <- shuffle[(n_A+1):length(all_values)]
    
    #Calculate dealta mu
    delta_mu <- mean(r_A) - mean(r_B)
    
    mu_vec[i] <- delta_mu
    
  }
  
  return(mu_vec)
  
}
```

## Linear Correlation Permuation Test

```{r, message = F, echo = T, message = F, eval = T}

perm_cor <- function(perms = 1000, x, y)
{
  ## Variables ##
  # perms: The number of permutations 
  # x: Vector of Variable 1 - for computing correlation
  # y: Vector of Variable 2 - for computing correlation
  ###############

  #Create zero vector
  vec <- numeric(perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    #Randomize y vector
    r_Y <- sample(y)
    
    #Computer correlation and store
    vec[i] <- cor(x, r_Y)
    
  }
  
  # Return new updated vector
  return(vec)
}
```

# Gapminder Questions

```{r, message = F, echo = T, message = F, eval = T}
#Calculating mean, and permutation test.
mean_diff <- mean(asia_vowels$rate_vowel) - mean(europe_vowels$rate_vowel)
vowel_test <- perm_mean(perms = 1000, vowels$rate_vowel, 33)

#Calculate p value.
p_percent <- 1 - (sum(vowel_test < mean_diff)/1000 * 100)
#As our p value is >5%, our result is not statistically significant
```

#Your data science team must come up with two specific questions of interest about the gapminder data set (from 'gapminder' library). 

Is there a correlation between life expectancy and population?

How does the number of vowels in a countries name differ between Europe and Asia?


#* Briefly describe the gapminder dataset.

This is a dataset compiled from UN collected data that tracks gdp, gdp per capita, population, and life expectancy across countries every five years. 

#* Explain why your questions are important. One of the questions should be answered with the two-sample mean permutation test, and the other with the correlation permutation test.

Our first question is important because it knowing how population and life expectancy are correlated can help decision makers to determine whether family planning is the best strategy to improve life expectancy. 

The second question is important because it provides entertainment to our domain expert.

#* Create visualizations of the statistical tests.
```{r, message = F, echo = T, message = F, eval = T}
ggplot() + aes(vowel_test)+ geom_histogram(binwidth=.01, colour="black", fill="white") + geom_vline(mapping =aes(xintercept=mean_diff))
```

#* Make data-based conclusions about the results. Do you believe there are differences among the groups you are comparing? 

#* Generate percentiles, summary statistics (max, min), and probabilites (p-values) to justify your conclusions.

#* Summarize each individual's contribution to the lab.
